---
title: "Effect of Polling Station Geometry on Voter Turnout"
author: "Serhii Tytiuk"
format: revealjs
---

## Problem

<font size="5">

-   Turnout in Ukraine elections is quite low compared to European countries

-   This is usually associated with political culture, distrust of elections, and so on.

-   However, there are many studies that show the dependence of turnout on the spatial characteristics of electoral units.

This small study will shed light on this aspect in the Ukrainian context </font>

## Dataset overview

<font size="5">The data was scraped before the full-scale war from the State Voters Register (polygons of electoral precincts) and the Central Election Commission (data on the 1st and 2nd rounds of 2019 Presidential elections)</font>

```{r}
#| warning: false
#| echo: false
#| output: false

library(raster)
library(sf)
library(terra)
library(tidyverse)
library(Rcpp)
library(tmap)
library(readxl)
library(stargazer)

#data import
shp <- '~/Desktop/kse_school/school_project/dilnytsi/dilnytsi.shp'
ps <- st_read(shp) %>%
  rename(ps_id = name)

territories <- read_excel("~/Desktop/kse_school/school_project/Polling_stations.xlsx", sheet = "PS") %>%
  rename(ps_id = "№ ВД", n_registered = "2.К-сть виборців, внесених до списку", region = 'Регіон',
         size = "size...16") %>%
  select(ps_id, region, oblast_y, size, cord_X, cord_Y, type)

elections_t1 <- read_delim("~/Desktop/kse_school/school_project/elections_t1.tsv", 
                           delim = "\t", escape_double = FALSE, 
                           trim_ws = TRUE)

round1 <- elections_t1 %>%
  rename(ps_id = "№ ВД", n_registered_1st = "2.К-сть виборців, внесених до списку",
         voters_1st = "9.К-сть виборців, які взяли участь у голосуванні") %>%
  select(ps_id, n_registered_1st, voters_1st) %>%
  mutate(turnout_1st = voters_1st/n_registered_1st, ps_id = as.double(ps_id))

elections_t2 <- read_delim("~/Desktop/kse_school/school_project/elections_t2.tsv", 
                           delim = "\t", escape_double = FALSE, 
                           trim_ws = TRUE)

round2 <- elections_t2 %>%
  rename(ps_id = "№ ВД", n_registered_2nd = "2.К-сть виборців, внесених до списку",
         voters_2nd = "9.К-сть виборців, які взяли участь у голосуванні") %>%
  select(ps_id, n_registered_2nd, voters_2nd) %>%
  mutate(turnout_2nd = voters_2nd/n_registered_2nd, ps_id = as.double(ps_id))

#merging the data
ps_mod <- ps %>%
  mutate(ps_id = as.double(ps_id)) %>%
  select(ps_id, geometry)

ps_full_round1 <- left_join(territories, round1, by="ps_id")
ps_full_round1 <- left_join(ps_full_round1, ps_mod, by="ps_id")
ps_full <- left_join(ps_full_round1, round2, by="ps_id")

#removing ps with invalid coordinates
bad_ps <- c(121606, 631539, 51675, 511452, 511453, 511267, 511272, 511451)

ps_full <- ps_full %>%
  filter(!(ps_id %in% bad_ps)) %>%
  mutate(area = as.numeric(st_area(geometry)/1000000), density = n_registered_1st/area)
```

<font size="5">

```{r results = 'asis'}
#| echo: false

sum_table <- ps_full %>% 
  select(n_registered_1st, voters_1st, n_registered_2nd, voters_2nd, turnout_1st, turnout_2nd, area, density)

stargazer(as.data.frame(sum_table), type = "html")
```

</font>

## Model

:::: {.columns}
::: {.column width="50%"}
<font size="3.5">

```{r results = "asis"}
#| warning: false
#| echo: false
model_dataset <- ps_full[sample(1:nrow(ps_full), 1000), ]
mod1 <- lm(turnout_1st ~ log(density) + log(n_registered_1st) + size + region, model_dataset)
mod2 <- lm(turnout_2nd ~ log(density) + log(n_registered_2nd) + size + region, model_dataset)
stargazer(mod1, mod2, type="html", single.row = TRUE)
```

</font>
:::

::: {.column width="50%"}
<font size="5"> As we can see, density as a proxy variable of precinct shape is a highly significant in both cases
</font>
:::
::::

## Mapping

```{r}
#| warning: false
#| echo: false

ps_viz <- st_sf(ps_full)
ps_viz %>%
  tm_shape() + 
  tm_fill("turnout_1st",
          palette = "PuBu",
          id="ps_id",
          popup.vars=c("turnout_1st")
  ) + 
  tm_legend(outside=TRUE) +
  tm_layout(frame = FALSE) 
```